Resolving the measurement uncertainty paradox in ecological management

### 2018-11-22 manuscripts/manuscript.tex
rebuild manuscript

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Resolving the measurement uncertainty paradox in ecological management},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

The challenge of making decisions under uncertainty is obviously not
unique to fisheries. Existing methods we optimal control methods which
similarly assume perfect measurements can be found in a wide range of
population and ecosystem management challenges, including fire ecology
(Richards et al. 1999), invasive species (Shea 1998; Blackwood et al.
2010), disease outbreaks (Shea et al. 2014; Li et al. 2017), and
protecting biodiversity (Dee et al. 2017; Iacona et al. 2017), and in
common textbooks and reviews (Mangel 1985; Clark 1990; Marescot et al.
2013). Yet fisheries conservation and management has long been both a
crucible and proving ground for the theory of ecological management more
generally, including topics such as adaptive management (Walters and
Hilborn 1978), ecosystem-based management (Levin and Lubchenco 2008) and
resilience thinking (Holling 1973; May 1977), while also giving rise to
the sub-discipline of resource economics (Gordon and Press 1954;
Schaefer 1954; Beverton and Holt 1957), thanks to its global relevance,
long history, and readily available data. Despite the generality of the
models and issue of measurement uncertainty considered here, any
application must be grounded in both the history and context of the
particular decision problem. In light of this, we begin with a brief
background on existing decision theory approaches common in fisheries
management.

In this paper, we examine three strategies to managing a natural
population, such as a fishery, under such uncertainty. This problem has
a long history in both theory and practice, where seemingly innocuous
simplifying assumptions have led to a deep paradox in how managers
should respond to uncertainty. The precautionary principle (Kriebel et
al. 2001) would seem to suggest that the greater the uncertainty in our
models, the more cautious we should be in management policies. Yet a
well-known mathematical proof by Reed (1979) established that under
quite general conditions, the optimal strategy under uncertainty is no
different than under no uncertainty at all. Attempts to resolve this
have since only deepened this paradox, arguing that adding additional
sources of uncertainty, such as imperfect measurements, should lead to
optimal harvest that increases with greater uncertainty (e.g. Clark and
Kirkwood 1986; Sethi et al. 2005). Meanwhile, much of actual practice
has (perhaps fortunately, as we shall see) ignored these results in
favor of a more heuristic application of the precautionary principle.
Here, we show that this paradox can be resolved by uncovering previous
shortcuts and tackling the optimal management problem head-on, which we
can solve thanks to the computational efficiency of sophisticated
algorithms developed more recently in the field of robotics. This
approach allows us to reject this paradox largely in favor of the
precautionary principle. Further, our comparisons show that widely
recognized existing approaches in both theory and practice are not by
themselves sufficient to manage a population under substantial
uncertainty over the long term, but our recently borrowed approach from
robotics literature can under identical conditions sustain both higher
biomass and higher economic yields than those approaches. This
conclusion has important implications not only for fisheries but
ecosystem more broadly: underscoring the danger of not correctly
accounting for measurement uncertainty when making decisions, and
further to the significant opportunity to learn from fields such as
robotics when faced with complex and uncertain decisions.

@@

where the \(\min\) function ensures no additional reward for attempting
to harvest more than the entire stock. In economic terms, the
corresponds to a fixed price per fish caught, with no proportional cost
on effort required to do so, but the results do not depend on this
assumption, as existing analyses have previously established (Reed
1979). Using this form allows a more direct comparison to approaches
such as maximum sustainable yield (MSY) which do not include an economic
model but merely maximize long-term catch. In principle, utility could
also reflect ecosystem services or other intrinsic values for fish left
in the sea and not merely the market for consumption (Halpern et al.
2013).

The solution to this decision problem is conditional on a population
model for the underlying ecological process, describing probability the
system moves to state \(x_{t+1}\) given that it was in state \(x_t\) and
the manager selected a harvest \(h_t\), which we discuss below. Such a
model would typically be estimated and parameterized from available
data, often (but not necessarily) using a hierarchical Bayesian approach
that acknowledges the existence of uncertainty in both measurements and
process (e.g.~see Dichmont et al. 2016 for a discussion of how
uncertainties are frequently included in stock assessment models).
However, accounting for measurement uncertainty when going from data to
model does not mean we have also accounted for this uncertainty when
going from model to decision. For example, a harvest decision policy
based upon maximum sustainable yield simply depends on the stock size
which maximizes the expected growth rate of the model. Measurement
uncertainty in model estimation can indeed influence the position of
that maximum, but as we shall see, that strategy is only optimal in the
absence of any uncertainty.

There is one further subtlety arising in this seemingly simple problem
statement that we must address. Given our discrete-time formulation of
the recruitment process, it is necessary to decide if the measurement
\(y_t\) happens before or after harvest \(h_t\): that is do we: measure,
recruit, harvest, or measure, harvest, recruit? Following convention in
the optimal control literature (e.g. (Reed 1979; Clark 1990; Sethi et
al. 2005)), we will assume the latter; measurement occurs immediately
before harvest, followed by recruitment. This assumption has
occasionally been reversed to introduce some additional uncertainty from
stochastic recruitment immediately prior to setting the harvest quota
(e.g. Clark and Kirkwood 1986; Weitzman 2002), which appears to mimic
uncertainty in measurements but does not allow uncertainty to accumulate
since the measurement step still occurs without error.

@@

Gordon and Press (1954) and Schaefer (1954) independently showed in the
same year that the maximum sustainable yield for the model bearing their
names is achieved by reducing the stock to the population to the size at
which it obtains the maximum growth rate which is known as Biomass at
Maximum Sustainable Yield, \(B_{MSY}\). For the Gordon-Schaefer model
(and many others), this is achieved at \(B_{MSY} = K/2\). They observed
that fishing at a \emph{constant yield} (that is, an individual fish
mortality, \(F\), such that harvest \(H = F \cdot B\)) of
\(F_{MSY} := \tfrac{H_{MSY}}{B_{MSY}}\)) will eventually lead to a
population that converges to the biomass \(B_{MSY}\) and produces the
maximum sustainable harvest, \(H_{MSY} = r K / 4\) in this model. From
here, theory and practice diverge.

@@

In the face of a naturally fluctuating population, previous arguments
made by Gordon and Press (1954); Schaefer (1954) or Clark (1973) no
longer hold. As noted above, the optimal harvest of a population under a
model such as the stochastic Gordon Schaefer, Eq
\eqref{Gordon-Schaefer}, is an example of a Markov Decision Problem
(MDP) which must be solved with stochastic dynamic programming. Thanks
to a mathematical proof provided by Reed (1979), fisheries managers have
largely been able to bi-pass such computational effort. However, this
convenient result also opened the door to paradox of uncertainty that
would haunt ecological management for at least the next four decades.

@@

Algorithmic innovations from the robotics community have now made it
possible to numerically solve non-trivial POMDP problems. The need to
consider all past observations becomes particularly demanding given the
forward-looking nature of a decision problem. Not only must we consider
all possible sequences of future actions over all possible sequences of
future states, but we must add to that all possible sequences of future
measurements. The computational complexity involved has historically
limited POMDP applications to contexts with only a handful of possible
states and actions (e.g. Chadès et al. 2008, 2011; Fackler and Haight
2014; Fackler and Pacifici 2014), insufficient to capture fisheries
models appropriately. Meanwhile, methods to solve decision problems with
imperfect information have advanced steadily in the field of robotics
and artificial intelligence. Almost contemporaneous with Reed's work,
papers by Smallwood and Sondik (1973) and Sondik (1978) laid
mathematical foundations for algorithms that could efficiently solve
POMDP problems (e.g. Kaelbling et al. 1998; Pineau et al. 2003). More
recently, driven by demands in areas such as autonomous vehicle decision
making, newer point-based approximation algorithms such as SARSOP
(Kurniawati et al. 2008) have made it possible to solve POMDPs with 100s
of states and actions. We have adapt this algorithm for application to
the fisheries context, and provide an efficient implementation as a
preliminary R package (Boettiger et al. 2018) which can be used to
replicate and further explore results presented here. Details of the
analysis presented here including annotated code to reproduce and
further explore all of the following results is provided in the
supplementary material.

Precise definitions of both the (fully observed) MDPs and POMDPs (not to
be confused with Markov Process Models and Hidden Markov Process Models
-- methods for which estimate the model but do not involve the feedback
of any decision process) in an ecological context can be found in
Williams (2011), or in classic work from the engineering and robotics
community (e.g. Smallwood and Sondik 1973, and @Sondik1978). For
convenience, we summarize the POMDP definition and common notation here.
The POMDP problem is posed identically to that of the MDP problem, with
the addition of an observation process The POMDP problem for fisheries
question considered here can be summarized as follows:

\begin{itemize}
\item
  Transition function (state equation): \(T(x_t, x_{t+1}, a_t)\): the
  probability that a system is in state \(x_{t+1}\) at time \(t+1\)
  given that it began in state \(x_t\) at time \(t\) and the manager
  took action \(a_t\). In our context, this relationship is given by the
  Gordon-Schaefer stock recruitment function \(f\) with normally
  distributed growth uncertainty
  \(x_{t+1} \sim \mathcal{N}(f(x_t,a_t), \sigma_g)\), truncated at zero
  to exclude negative population sizes. The action \(a_t\) represents
  the harvest quota set for the fishery. Attempting to implement a quota
  the exceeds the true population size results in collapse of the
  fishery. As Clark and Kirkwood (1986) eloquently argues, assuming that
  extinction is impossible regardless of fishing intensity would
  unreasonably bias the decision problem.
\item
  Observation function: \(O(x_t,y_t,a_{t-1})\) the probability of
  observing state \(y_t\) given a system in state \(x_t\). In principle,
  the action chosen can influence the precision of the observation. In
  our case, we simply assume normally distributed errors around the true
  state, \(y_t \sim \mathcal{N}(x_t, \sigma_m)\), truncated at zero to
  exclude negative population sizes.
\item
  Utility function: \(U(x_t,a_t)\), the utility received at time \(t\)
  for taking action \(a_t\), given that the system is in state \(x_t\).
  For simplicity of analysis, we will simply set the utility to be equal
  to the harvested stock: \(U(x_t, a_t) = \min(x_t, a_t)\), indicating
  that realized harvest cannot be negative. This choice ensures that in
  the case of no uncertainty (\(\sigma_g = \sigma_m = 0\)), the optimal
  solution matches that expected under a simple MSY calculation. More
  realistic utility functions may include diminishing returns with
  increasing harvest (supply and demand effects), and the cost of
  fishing, both of which act to suppress large harvests. By focusing on
  a simple utility we can be sure that our comparison to MSY is driven
  by the treatment of uncertainty rather than merely differing economic
  assumptions.
\end{itemize}

@@

In the absence of either environmental noise or measurement error POMDP,
CE, and MSY would converge to stock at the \(B_{MSY}\), while TAC would
maintain the stock at a slightly higher level (see Supplemental
material, Figure S1). The introduction of stochastic growth has a
significant negative impact on the TAC strategy, with a mean biomass
significantly lower than \(B_{MSY}\) (first panel, Figure \ref{sims};
this impact is even more severe for MSY, see Figure S2). Without
measurement error, the CE and POMDP strategies are nearly identical with
both approximately maintaining the stock at \(B_{MSY}\) despite the
significant environmental stochasticity. As measurement error increases
in the subsequent panels, CE and TAC strategies perform increasingly
poorly, while the POMDP continues to maintain the average stock close to
\(B_{MSY}\). Notably, CE is significantly more impacted by measurement
uncertainty than TAC, with CE averaging even lower biomass than TAC
under moderate measurement uncertainty. These results confirm that the
precautionary approach represented by the TAC does indeed prove more
robust to the problem of measurement error than the optimization
solution represented by CE, but is less robust to environmental error.
In contrast, the POMDP approach successfully handles both sources of
uncertainty, maintaining higher stocks, near \(B_{MSY}\) level, and
producing the highest yield.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/econ-1.pdf}
\caption{Expected net present value of the fishery across strategies
under increasing levels of measurement error \(\sigma_m\). Net present
value is the sum of all future harvests discounted over time and
averaged across 100 replicates simulations under each strategy, as shown
in Figure \ref{sims}. The value under constant escapement (CE) is
optimal when measurements are perfect but decreases rapidly with
increasing measurement error. The more cautious TAC is not economically
optimal but largely unimpacted by increasing error, while POMDP attains
consistently high economic yield despite the increasing uncertainty.
\label{econ}}
\end{figure}

This pattern in management success in ecological terms (the relatively
recovery and maintenance of fish biomass) is also borne out in terms of
economic performance. We find the mean net present value (averaging
across replicates and discounting future profits by the discount factor
\(\gamma\)) for these same simulations at increasing levels of
measurement uncertainty (Figure \ref{econ} ). As before, environmental
stochasticity is set at \(\sigma_g = 0.15\). In the absence of
measurement uncertainty, constant escapement (CE) is optimal (as per
Reed (1979)), while the heuristic caution built into the total allowable
catch (TAC) strategy (at 80\% MSY) results in a sub-optimal economic
yield. The presence of stochasticity in recruitment
(\(\sigma_g = 0.15\)), also contributes to the reduction in yield under
TAC. Though CE is quite robust to the environmental stochasticity level,
this strategy proves very sensitive to increasing measurement error,
showing sharp declines in economic value, falling below the TAC economic
value at \(\sigma_m = 0.15\). Though the risk of stock collapse does
increase with increasing measurement error under TAC (as seen by the
mean declines in Figure \ref{sims}), these have little impact on the
economic value due to the discount rate. A smaller discount rate would
penalize unlikely but not improbable collapses more, since a significant
amount of time is required to realize those rare events. Supplemental
Figure S3 summarizes these economic trends across different noise values
\(\sigma_g\) and includes comparison to a simple MSY policy. In contrast
to the declining economic performance of CE and the consistently
sub-optimal economic yield for TAC, the POMDP strategy continues to
generate economic yields at approximately the optimal level (as attained
by CE in the absence of measurement error) despite the increasingly
uncertain measurements. This demonstrates that the reduced risk of stock
collapse and higher average stock biomass attained by the POMDP strategy
in the presence of high measurement error, Figure \ref{sims}) is not the
result of a trivial reduction in harvesting under all circumstances, but
rather, evidence of a more nuanced strategy that manages to account for
the uncertainty in measurement while maintaining a reasonable harvest.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/overest-1.pdf}
\caption{Average fish biomass when the POMDP strategy assumes a high
level of measurement uncertainty, while simulations reflect perfect
measurements. Despite this overestimation, biomass under POMDP strategy
closely tracks the optimal biomass under CE in the absence of
measurement error. \label{overest}}
\end{figure}

@@

This difference between underestimating and overestimating accounts for
the poor performance of the CE solution under large measurement
uncertainty, where it over-harvests whenever measurements are too large.
Even though underestimating is equally likely under the measurement
uncertainty model, sooner or later a run of ``heads'', a sequence of
overestimations relative to the true stock, can drive stocks to very low
levels where the chance of local extinction becomes possible. It is
precisely this asymmetry: that too much over-harvesting leads to an
irreversible state of extinction, while too much under-harvesting is
always reversible (modulo some lost revenue) that lay behind Clark's
original intuition that there was something fishy about Reed's result
that \(S = D\): that uncertainty required no extra caution. Constant
escapement is particularly susceptible to over-harvesting starting from
stock sizes much higher than \(B_{MSY}\) since its bang-bang optimal
solution attempts to bring stocks back to target level as fast as
possible. CE is not so vulnerable to collapse from small stock sizes,
since it shuts down all harvests once estimates fall below \(B_{MSY}\).
Stock collapse under TAC can also be driven by such a string of heads,
but is unlikely at high stock sizes since harvest never exceeds
\(H_{MSY}\). In both cases, some measurement uncertainty interacts with
inherent stochasticity, which provides a continued source of variation
to sizes above \(B_{MSY}\), where CE strategy is most vulnerable, and
below \(B_{MSY}\), where TAC is most vulnerable.

>>>>

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Resolving the measurement uncertainty paradox in ecological management},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

The challenge of making decisions under uncertainty is not unique to
fisheries. Optimal control-based decision methods which assume perfect
measurements are applied to a wide range of population and ecosystem
management issues, including fire ecology (Richards et al. 1999),
invasive species (Shea 1998; Blackwood et al. 2010), disease outbreaks
(Shea et al. 2014; Li et al. 2017), and protecting biodiversity (Dee et
al. 2017; Iacona et al. 2017), and in common textbooks and reviews
(Mangel 1985; Clark 1990; Marescot et al. 2013). Yet fisheries
conservation and management has long been both a crucible and proving
ground for the theory of ecological management more generally, including
topics such as adaptive management (Walters and Hilborn 1978),
ecosystem-based management (Levin and Lubchenco 2008) and resilience
thinking (Holling 1973; May 1977), while also giving rise to the
sub-discipline of resource economics (Gordon and Press 1954; Schaefer
1954; Beverton and Holt 1957), thanks to its global relevance, long
history, and readily available data. Despite the generality of the
models and issue of measurement uncertainty considered here, any
application must be grounded in both the history and context of the
particular decision problem. In light of this, we begin with a brief
background on existing decision theory approaches common in fisheries
management.

In this paper, we examine three strategies to managing a natural
population, such as a fishery, under such uncertainty. This problem has
a long history in both theory and practice, where seemingly innocuous
simplifying assumptions have led to a deep paradox in how managers
should respond to uncertainty. The precautionary principle (Kriebel et
al. 2001) would seem to suggest that the greater the uncertainty in our
models, the more cautious we should be in management policies. Yet a
well-known mathematical proof by Reed (1979) established that under
quite general conditions, the optimal strategy under uncertainty is no
different than under no uncertainty at all. Attempts to resolve this
have since only deepened this paradox, arguing that adding additional
sources of uncertainty, such as imperfect measurements, should lead to
optimal harvest that increases with greater uncertainty (e.g. Clark and
Kirkwood 1986; Sethi et al. 2005). Meanwhile, much of actual practice
has (perhaps fortunately, as we shall see) ignored these results in
favor of a more heuristic application of the precautionary principle.
Here, we show that this paradox can be resolved by uncovering previous
shortcuts and tackling the optimal management problem head-on, which we
can solve thanks to the computational efficiency of sophisticated
algorithms developed more recently in the field of robotics. This
approach allows us to reject this paradox largely in favor of the
precautionary principle. Further, our comparisons show that widely
recognized existing approaches in both theory and practice are not by
themselves sufficient to manage a population under substantial
uncertainty over the long term, but our recently borrowed approach from
robotics literature can under identical conditions sustain both higher
biomass and higher economic yields than those approaches. This
conclusion has important implications not only for fisheries but
ecosystem more broadly: First, our results highlight that measurement
error must be accounted for in both the decision process itself, and not
merely in the estimation of model. Second, these results underscore the
opportunity to learn from engineering fields such as robotics when faced
with complex and uncertain decisions that we have previously dismissed
as intractable.

@@

where the \(\min\) function ensures no additional reward for attempting
to harvest more than the entire stock. In economic terms, this
corresponds to a fixed price per fish caught, with no proportional cost
on effort required to do so, but the results do not depend on this
assumption, as existing analyses have previously established (Reed
1979). Using this form allows a more direct comparison to approaches
such as maximum sustainable yield (MSY) which do not include an economic
model but merely maximize long-term catch. In principle, utility could
also reflect ecosystem services or other intrinsic values for fish left
in the sea and not merely the market for consumption (Halpern et al.
2013).

The solution to this decision problem is conditional on a population
model for the underlying ecological process, describing the probability
that the system moves to state \(x_{t+1}\) given that it was in state
\(x_t\) and the manager selected a harvest \(h_t\), which we discuss
below. Such a model would typically be estimated and parameterized from
available data, often (but not necessarily) using a hierarchical
Bayesian approach that acknowledges the existence of uncertainty in both
measurements and process (e.g.~see Dichmont et al. 2016 for a discussion
of how uncertainties are frequently included in stock assessment
models). However, accounting for measurement uncertainty when going from
data to model does not mean we have also accounted for this uncertainty
when going from model to decision. For example, a harvest decision
policy based upon maximum sustainable yield simply depends on the stock
size which maximizes the expected growth rate of the model. Measurement
uncertainty in model estimation can indeed influence the position of
that maximum, but as we shall see, that strategy is only optimal in the
absence of any uncertainty.

There is one further subtlety arising in this seemingly simple problem
statement that we must address. Given our discrete-time formulation of
the recruitment process, it is necessary to decide if the measurement
\(y_t\) happens before or after harvest \(h_t\): that is do we: measure,
recruit, harvest, or measure, harvest, recruit? Following convention in
the optimal control literature (e.g. Reed 1979; Clark 1990; Sethi et al.
2005), we will assume the latter; measurement occurs immediately before
harvest, followed by recruitment. This assumption has occasionally been
reversed to introduce some additional uncertainty from stochastic
recruitment immediately prior to setting the harvest quota (e.g. Clark
and Kirkwood 1986; Weitzman 2002), which appears to mimic uncertainty in
measurements but does not allow uncertainty to accumulate since the
measurement step still occurs without error.

@@

Gordon and Press (1954) and Schaefer (1954) independently showed that
the maximum sustainable yield is achieved by reducing the stock to the
population to the size at which it obtains the maximum growth rate. This
stock size is known as Biomass at Maximum Sustainable Yield,
\(B_{MSY}\). For the Gordon-Schaefer model (and many others), this is
achieved at \(B_{MSY} = K/2\). They observed that fishing at a
\emph{constant yield} (that is, an individual fish mortality, \(F\),
such that harvest \(H = F \cdot B\)) of
\(F_{MSY} := \tfrac{H_{MSY}}{B_{MSY}}\)) will eventually lead to a
population that converges to the biomass \(B_{MSY}\) and produces the
maximum sustainable harvest, \(H_{MSY} = r K / 4\) in this model. From
here, theory and practice diverge.

@@

In the face of a naturally fluctuating population, previous arguments
made by Gordon and Press (1954); Schaefer (1954) or Clark (1973) no
longer hold. As noted above, the optimal harvest of a population under a
model such as the stochastic Gordon Schaefer, Eq
\eqref{Gordon-Schaefer}, is an example of a Markov Decision Problem
(MDP) which must be solved with stochastic dynamic programming. Thanks
to a mathematical proof provided by Reed (1979), fisheries managers have
largely been able to bi-pass such computational effort. However, this
convenient result also opened the door to paradox of uncertainty that
has persisted in ecological management for the past four decades.

@@

Algorithmic innovations from the robotics community have now made it
possible to numerically solve non-trivial POMDP problems. The need to
consider all past observations becomes particularly demanding given the
forward-looking nature of a decision problem. Not only must we consider
all possible sequences of future actions over all possible sequences of
future states, but we must add to that all possible sequences of future
measurements. The computational complexity involved has historically
limited POMDP applications to contexts with only a handful of possible
states and actions (e.g. Chadès et al. 2008, 2011; Fackler and Haight
2014; Fackler and Pacifici 2014), insufficient to capture fisheries
models appropriately. Meanwhile, methods to solve decision problems with
imperfect information have advanced steadily in the field of robotics
and artificial intelligence. Almost contemporaneous with Reed's work,
papers by Smallwood and Sondik (1973) and Sondik (1978) laid
mathematical foundations for algorithms that could efficiently solve
POMDP problems (e.g. Kaelbling et al. 1998; Pineau et al. 2003). More
recently, driven by demands in areas such as autonomous vehicle decision
making, newer point-based approximation algorithms such as SARSOP
(Kurniawati et al. 2008) have made it possible to solve POMDPs with 100s
of states and actions. We have adapted this algorithm for application to
the fisheries context, and provide an efficient implementation as a
preliminary R package (Boettiger et al. 2018) which can be used to
replicate and further explore results presented here. Details of the
analysis presented here including annotated code to reproduce and
further explore all of the following results are provided in the
supplementary material.

Precise definitions of both the (fully observed) MDPs and POMDPs (not to
be confused with Markov Process Models and Hidden Markov Process Models
-- methods for which estimate the model but do not involve the feedback
of any decision process) in an ecological context can be found in
Williams (2011), or in classic work from the engineering and robotics
community (e.g. Smallwood and Sondik 1973; Sondik 1978). For
convenience, we summarize the POMDP definition and common notation here.
The POMDP problem is posed identically to that of the MDP problem, with
the addition of an observation process The POMDP problem for fisheries
question considered here can be summarized as follows:

\begin{itemize}
\item
  Transition process (state equation): \(T(x_t, x_{t+1}, a_t)\): the
  probability that a system is in state \(x_{t+1}\) at time \(t+1\)
  given that it began in state \(x_t\) at time \(t\) and the manager
  took action \(a_t\). In our context, this relationship is given by the
  Gordon-Schaefer stock recruitment function \(f\) with normally
  distributed growth uncertainty
  \(x_{t+1} \sim \mathcal{N}(f(x_t,a_t), \sigma_g)\), truncated at zero
  to exclude negative population sizes. The action \(a_t\) represents
  the harvest quota set for the fishery. Attempting to implement a quota
  the exceeds the true population size results in collapse of the
  fishery. As Clark and Kirkwood (1986) eloquently argues, assuming that
  extinction is impossible regardless of fishing intensity would
  unreasonably bias the decision problem.
\item
  Observation function: \(O(x_t,y_t,a_{t-1})\) the probability of
  observing state \(y_t\) given a system in state \(x_t\). In principle,
  the action chosen can influence the precision of the observation. In
  our case, we simply assume normally distributed errors around the true
  state, \(y_t \sim \mathcal{N}(x_t, \sigma_m)\), truncated at zero to
  exclude negative population sizes.
\item
  Utility function: \(U(x_t,a_t)\), the utility received at time \(t\)
  for taking action \(a_t\), given that the system is in state \(x_t\).
  For simplicity of analysis, we will simply set the utility to be equal
  to the harvested stock: \(U(x_t, a_t) = \min(x_t, a_t)\), indicating
  that realized harvest cannot be negative. This choice ensures that in
  the case of no uncertainty (\(\sigma_g = \sigma_m = 0\)), the optimal
  solution matches that expected under a simple MSY calculation. More
  realistic utility functions may include diminishing returns with
  increasing harvest (supply and demand effects), and the cost of
  fishing, both of which act to suppress large harvests. By focusing on
  a simple utility we can be sure that our comparison to MSY is driven
  by the treatment of uncertainty rather than merely differing economic
  assumptions.
\end{itemize}

@@

In the absence of either environmental noise or measurement error POMDP,
CE, and MSY would converge to stock at the \(B_{MSY}\), while TAC would
maintain the stock at a slightly higher level (see Supplemental
material, Figure S1). The introduction of stochastic growth has a
significant negative impact on the TAC strategy, with a mean biomass
significantly lower than \(B_{MSY}\) (first panel, Figure \ref{sims};
this impact is even more severe for MSY, see Figure S2). Without
measurement error, the CE and POMDP strategies are nearly identical with
both approximately maintaining the stock at \(B_{MSY}\) despite the
significant environmental stochasticity. As measurement error increases
in the subsequent panels, CE and TAC strategies perform increasingly
poorly, while the POMDP continues to maintain the average stock close to
\(B_{MSY}\). Notably, CE is significantly more impacted by measurement
uncertainty than TAC, with CE averaging even lower biomass than TAC
under moderate measurement uncertainty. These results confirm that the
precautionary approach represented by the TAC does indeed prove more
robust to the problem of measurement error than the optimization
solution represented by CE, but is less robust to environmental error.
Many of the replicate simulations under CE experience complete stock
extinction under large measurement error, owing to the arbitrarily large
harvests permitted by the ``bang-bang'' nature of the CE policy. In
contrast, the POMDP approach successfully handles both sources of
uncertainty, maintaining higher stocks, near \(B_{MSY}\) level, and
producing the highest yield.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/econ-1.pdf}
\caption{Expected net present value of the fishery across strategies
under increasing levels of measurement error \(\sigma_m\). Net present
value is the sum of all future harvests discounted over time and
averaged across 100 replicates simulations under each strategy, assuming
discount factor \(\gamma = 0.95\), as shown in Figure \ref{sims}. The
value under constant escapement (CE) is optimal when measurements are
perfect but decreases rapidly with increasing measurement error. The
more cautious TAC is not economically optimal but largely unimpacted by
increasing error, while POMDP attains consistently high economic yield
despite the increasing uncertainty. \label{econ}}
\end{figure}

This pattern in management success in ecological terms (the relatively
recovery and maintenance of fish biomass) is also borne out in terms of
economic performance. We find the mean net present value (averaging
across replicates and discounting future profits by the discount factor
\(\gamma = 0.95\)) for these same simulations at increasing levels of
measurement uncertainty (Figure \ref{econ}). As before, environmental
stochasticity is set at \(\sigma_g = 0.15\). In the absence of
measurement uncertainty, constant escapement (CE) is optimal (as per
Reed (1979)), while the heuristic caution built into the total allowable
catch (TAC) strategy (at 80\% MSY) results in a sub-optimal economic
yield. The presence of stochasticity in recruitment
(\(\sigma_g = 0.15\)), also contributes to the reduction in yield under
TAC. Though CE is quite robust to the environmental stochasticity level,
this strategy proves very sensitive to increasing measurement error,
showing sharp declines in economic value, falling below the TAC economic
value at \(\sigma_m = 0.15\). Though the risk of stock collapse does
increase with increasing measurement error under TAC (as seen by the
mean declines in Figure \ref{sims}), these have little impact on the
economic value due to the discount rate. A smaller discount rate would
penalize unlikely but not improbable collapses more, since a significant
amount of time is required to realize those rare events. Supplemental
Figure S3 summarizes these economic trends across different noise values
\(\sigma_g\) and includes comparison to a simple MSY policy. In contrast
to the declining economic performance of CE and the consistently
sub-optimal economic yield for TAC, the POMDP strategy continues to
generate economic yields at approximately the optimal level (as attained
by CE in the absence of measurement error) despite the increasingly
uncertain measurements. This demonstrates that the reduced risk of stock
collapse and higher average stock biomass attained by the POMDP strategy
in the presence of high measurement error, Figure \ref{sims} is not the
result of a trivial reduction in harvesting under all circumstances, but
rather, evidence of a more nuanced strategy that manages to account for
the uncertainty in measurement while maintaining a reasonable harvest.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/overest-1.pdf}
\caption{Average fish biomass when the POMDP strategy assumes a high
level of measurement uncertainty, while simulations reflect perfect
measurements. Despite this overestimation, biomass under POMDP strategy
closely tracks the optimal biomass under CE in the absence of
measurement error. Solid lines indicate averages over 100 simulations,
colored bands indicate +/- one standard deviation. \label{overest}}
\end{figure}

@@

This difference between underestimating and overestimating accounts for
the poor performance of the CE solution under large measurement
uncertainty, where it over-harvests whenever measurements are too large.
Even though underestimating is equally likely under the measurement
uncertainty model, sooner or later a run of ``heads'', a sequence of
overestimations relative to the true stock, can drive stocks to very low
levels where the chance of stock extinction becomes possible. It is
precisely this asymmetry: that too much over-harvesting leads to an
irreversible state of extinction, while too much under-harvesting is
always reversible (modulo some lost revenue) that lay behind Clark's
original intuition that there was something fishy about Reed's result
that \(S = D\): that uncertainty required no extra caution. Constant
escapement is particularly susceptible to over-harvesting starting from
stock sizes much higher than \(B_{MSY}\) since its bang-bang optimal
solution attempts to bring stocks back to target level as fast as
possible. CE is not so vulnerable to collapse from small stock sizes,
since it shuts down all harvests once estimates fall below \(B_{MSY}\).
Stock collapse under TAC can also be driven by such a string of heads,
but is unlikely at high stock sizes since harvest never exceeds
\(H_{MSY}\). In both cases, some measurement uncertainty interacts with
inherent stochasticity, which provides a continued source of variation
to sizes above \(B_{MSY}\), where CE strategy is most vulnerable, and
below \(B_{MSY}\), where TAC is most vulnerable.
